Quantizing the Bosonic String on a Loop Quantum Gravity Background

### 2024-06-21 lqg-strings.tex
moved emergent dimensions section into main text, made changes where approppriate and added a missing word in the introduction

In the decades prior to the advent of string theory a great deal of effort was put into attempting to quantize gravity via both the Lagrangian and Hamiltonian approaches. The Hamiltonian approach involves, as stated earlier, a choice of the foliation of the background geometry into a collection of spacelike surfaces $\Sigma_t$. Central to this approach was the ADM (Arnowitt-Deser-Misner) formalism which allows one to construct a Hamiltonian $H_{GR}$ for general relativity starting from the Einstein-Hilbert action $S_{EH}$. The resulting phase is co-ordinatized by a configuration variable which is the 3-metric $h_{ab}$ of the ``leaves'' $\Sigma_t$ of the foliation and a momentum variable which is a function of a the extrinsic curvature $k_{ab}$ of the leaves\footnote{Our notational convention is the following. Lowercase Greek letters from the middle of the alphabet $\mu, \nu, \ldots$ are spacetime indices which run from $(0 \ldots 3)$, whereas lowercase Roman letters $a,b,\ldots$ are spatial indices for quantities which live solely on the surfaces $\Sigma_t$ and take values in $(1,2,3)$}. The gravity Hamiltonian turns out to be a sum of two constraints known as the diffeomorphism (or ``momentum'') constraint $H_{diff}$ and the Hamiltonian constraint $H_{ham}$:
\begin{equation}
	H_{GR} = H_{diff} + H_{ham},
\end{equation}
and these are, in turn, functionals on the phase space of general relativity written in terms of $h_{ab}$ and $k_{ab}$. The idea then is that physical states of the theory $\ket{\Psi_{phys}} $, which can be written as functionals $\Psi[h_{ab}]$ of the 3-metric $h_{ab}$ , must be annihilated by these constraints:
\begin{equation}
	H_{GR} \ket{\Psi_{phys}} \equiv 0
\end{equation}
While this procedure is straightforward in principle, in practice is was impossible to implement in the quantum theory due to highly complicated non-polynomial dependence of the diffeomorphism and Hamiltonian constraints on the configuration and momentum variables. This unfortunate state of affairs persisted until the late 1980s when Abhay Ashtekar \cite{Ashketar1986New,Ashtekar1987New} recast general relativity as a theory of a Minkowski tetrad $E^I_\mu$ and a self-dual $\sltwoc$ connection $A_\mu^{IJ}$. Ashtekar realised that the constraints of general relativity, when expressed in terms of these ``new variables'', simplified drastically and became polynomial functions of the tetrad and connection variables. It was quickly realised by various researchers that the resulting theory could be quantized using the same methods used to quantize Yang-Mills gauge theory \cite{Ashtekar1992Weaving,Ashtekar2004Background} and that the diffeomorphism constraint could be solved exactly in terms of so-called \emph{spin-network} states \cite{Rovelli1988Knot,Rovelli1990Loop,Ashtekar1992Gravitons}.

@@

As for variation \wrt the embbeding fields $X^{\lambda}$, we have
\begin{equation} \label{eqn:vary-wrt-X}
	\frac{\delta S_{I A P}}{\delta X^{\lambda}} = 0
\end{equation}
\begin{multline} \label{eqn:X-eom-inverse-area}
	F^{-} \sqrt{- g} g^{a b} \partial _{a} \partial _{b} X^{\nu} G_{\lambda \nu} + \partial _{a} \left( F^{-} \sqrt{- g} g^{a b} \right) \partial _{b} X^{\nu} G_{\lambda \nu} = \\ = \frac{1}{2} F^{-} \sqrt{- g} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \partial _{\lambda} G_{\mu \nu} ,
\end{multline}
where
\begin{equation} \label{eqn:F-minus-def}
	F^{-} \equiv 1 - \frac{k \Delta}{2 (- g)} .
\end{equation}
Since all 2-dimensional metrics are conformally flat \cite{Schottenloher}, we use conformal gauge $g_{a b} = \e ^{\phi (x)} \eta _{a b}$ and assume flat background $G_{\mu \nu} = \eta _{\mu \nu}$, to simplify eq (\ref{eqn:X-eom-inverse-area}) into
\begin{equation} \label{eqn:X-eom-simplified}
	\eta ^{a b} \partial _{a} \partial _{b} X^{\mu} + \frac{1}{F^{-}} \eta ^{a b} \partial _{a} F^{-} \partial _{b} X^{\mu} = 0 .
\end{equation}
By expanding the definition of $F^{-}$ we get
\begin{equation} \label{eqn:coupling-X-phi}
	\eta ^{a b} \partial _{a} \partial _{b} X^{\mu} + \eta ^{a b} \frac{\left( 2 \del _{a} \phi - \del _{a} \ln (\Delta) \right)}{\dfrac{2 \e ^{2 \phi}}{k \Delta} - 1} \del _{b} X^{\mu} = 0 ,
\end{equation}
revealing a wave equation sourced by a coupling between the WS conformal factor, quantum of area field $\Delta$ and the embedding fields much akin to the dilaton field \cite{Balthazar-t-backgrounds}. We'll come back to this in \autoref{sec:discussion}.

@@

@@

In section \ref{sec:quantum-KG-string} we saw that our model allows closed string spin-1 excitations at the first level, which is due to our solution not having independent left and right sectors. Closed string excitations are usually related to the gravitational field and background spacetime itself, so what could the first level vector states of the closed string relate to? Well, in section \ref{sec:gauging-lt} we discussed about a possible formulation of ST in terms of a Lorentz connection $\mc{A}_{\mu}$, which is an ingredient in describing the gravitational field in tetrad-connection \cite{Bodendorfer-LQG} formulation, and so this seems to be the natural candidate as to what the first level vector excitation of the closed string could be,
\begin{equation} \label{eqn:connection-string-vector-excitation}
	\mc{A}_{\mu} \sim \eta _{\mu \nu} (\Hat{a}^{\nu} _{1})^{\dagger} \ket{0} .
\end{equation}
Another candidate for what the vector excitations could relate to is the Ashtekar electric fields $\widetilde{E}^{\mu} _{I}$,
\begin{equation} \label{eqn:AEF-string-vector-excitation}
	\widetilde{E}^{\mu} _{I} \sim (\Hat{a}^{\mu} _{I})^{\dagger} \ket{0} .
\end{equation}

@@

\subsection{Emergent Dimensions and LQG} \label{sec:emergent-dimensions}

\begin{acknowledgments}

\end{acknowledgments}

>>>>

In the decades prior to the advent of string theory a great deal of effort was put into attempting to quantize gravity via both the Lagrangian and Hamiltonian approaches. The Hamiltonian approach involves, as stated earlier, a choice of the foliation of the background geometry into a collection of spacelike surfaces $\Sigma_t$. Central to this approach was the ADM (Arnowitt-Deser-Misner) formalism which allows one to construct a Hamiltonian $H_{GR}$ for general relativity starting from the Einstein-Hilbert action $S_{EH}$. The resulting phase space is co-ordinatized by a configuration variable which is the 3-metric $h_{ab}$ of the ``leaves'' $\Sigma_t$ of the foliation and a momentum variable which is a function of a the extrinsic curvature $k_{ab}$ of the leaves\footnote{Our notational convention is the following. Lowercase Greek letters from the middle of the alphabet $\mu, \nu, \ldots$ are spacetime indices which run from $(0 \ldots 3)$, whereas lowercase Roman letters $a,b,\ldots$ are spatial indices for quantities which live solely on the surfaces $\Sigma_t$ and take values in $(1,2,3)$}. The gravity Hamiltonian turns out to be a sum of two constraints known as the diffeomorphism (or ``momentum'') constraint $H_{diff}$ and the Hamiltonian constraint $H_{ham}$:
\begin{equation}
	H_{GR} = H_{diff} + H_{ham},
\end{equation}
and these are, in turn, functionals on the phase space of general relativity written in terms of $h_{ab}$ and $k_{ab}$. The idea then is that physical states of the theory $\ket{\Psi_{phys}} $, which can be written as functionals $\Psi[h_{ab}]$ of the 3-metric $h_{ab}$ , must be annihilated by these constraints:
\begin{equation}
	H_{GR} \ket{\Psi_{phys}} \equiv 0
\end{equation}
While this procedure is straightforward in principle, in practice is was impossible to implement in the quantum theory due to highly complicated non-polynomial dependence of the diffeomorphism and Hamiltonian constraints on the configuration and momentum variables. This unfortunate state of affairs persisted until the late 1980s when Abhay Ashtekar \cite{Ashketar1986New,Ashtekar1987New} recast general relativity as a theory of a Minkowski tetrad $E^I_\mu$ and a self-dual $\sltwoc$ connection $A_\mu^{IJ}$. Ashtekar realised that the constraints of general relativity, when expressed in terms of these ``new variables'', simplified drastically and became polynomial functions of the tetrad and connection variables. It was quickly realised by various researchers that the resulting theory could be quantized using the same methods used to quantize Yang-Mills gauge theory \cite{Ashtekar1992Weaving,Ashtekar2004Background} and that the diffeomorphism constraint could be solved exactly in terms of so-called \emph{spin-network} states \cite{Rovelli1988Knot,Rovelli1990Loop,Ashtekar1992Gravitons}.

@@

As for variation \wrt the embbeding fields $X^{\lambda}$, we have
\begin{equation} \label{eqn:vary-wrt-X}
	\frac{\delta S_{I A P}}{\delta X^{\lambda}} = 0
\end{equation}
\begin{multline} \label{eqn:X-eom-inverse-area}
	F^{-} \sqrt{- g} g^{a b} \partial _{a} \partial _{b} X^{\nu} G_{\lambda \nu} + \partial _{a} \left( F^{-} \sqrt{- g} g^{a b} \right) \partial _{b} X^{\nu} G_{\lambda \nu} = \\ = \frac{1}{2} F^{-} \sqrt{- g} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \partial _{\lambda} G_{\mu \nu} ,
\end{multline}
where
\begin{equation} \label{eqn:F-minus-def}
	F^{-} \equiv 1 - \frac{k \Delta}{2 (- g)} .
\end{equation}
Since all 2-dimensional metrics are conformally flat \cite{Schottenloher}, we use conformal gauge $g_{a b} = \e ^{\phi (x)} \eta _{a b}$ and assume flat background $G_{\mu \nu} = \eta _{\mu \nu}$, to simplify eq (\ref{eqn:X-eom-inverse-area}) into
\begin{equation} \label{eqn:X-eom-simplified}
	\eta ^{a b} \partial _{a} \partial _{b} X^{\mu} + \frac{1}{F^{-}} \eta ^{a b} \partial _{a} F^{-} \partial _{b} X^{\mu} = 0 .
\end{equation}
By expanding the definition of $F^{-}$ we get
\begin{equation} \label{eqn:coupling-X-phi}
	\eta ^{a b} \partial _{a} \partial _{b} X^{\mu} + \eta ^{a b} \frac{\left( 2 \del _{a} \phi - \del _{a} \ln (\Delta) \right)}{\dfrac{2 \e ^{2 \phi}}{k \Delta} - 1} \del _{b} X^{\mu} = 0 ,
\end{equation}
revealing a wave equation sourced by a coupling between the WS conformal factor, quantum of area field $\Delta$ and the embedding fields much akin to the dilaton field \cite{Balthazar-t-backgrounds}. We'll come back to this in \autoref{sec:emergent-dimensions}.

@@

\section{Emergent Dimensions and LQG} \label{sec:emergent-dimensions}

@@

In section \ref{sec:quantum-kg-string} we saw that our area-corrected model allows closed string spin-1 excitations, which is due to our solution not having distinct left and right sectors and the physical states not being invariant under rigid rotations of the string. Closed string excitations are usually related to the gravitational field and background spacetime itself, so what could the vector states of the closed string relate to? Well, in section \ref{sec:gauging-lt} we discussed a possible formulation of ST in terms of a Poincaré connection $\mc{A}_{\mu}$, which is an ingredient in describing the gravitational field in tetrad-connection \cite{Bodendorfer-LQG} formulation, and so this seems to be the natural candidate as to what the first level vector excitation of the closed string could be,
\begin{equation} \label{eqn:connection-string-vector-excitation}
	\mc{A}_{\mu} \sim \eta _{\mu \nu} (\Hat{a}^{\nu} _{1})^{\dagger} \ket{0} .
\end{equation}
Another candidate for what the vector excitations could relate to is the Ashtekar electric fields $\widetilde{E}^{\mu} _{I}$,
\begin{equation} \label{eqn:AEF-string-vector-excitation}
	\widetilde{E}^{\mu} _{I} \sim (\Hat{a}^{\mu} _{I})^{\dagger} \ket{0} .
\end{equation}

@@

\begin{acknowledgments}
	DV would like to thank Thomas van Riet for going over an early draft of this work and for his valuable comments. DV and LdS would both like to thank David Schmidtt for his very valuable observations regarding the nature of the $\Delta$ field and its relation to the Kalb-Ramond field. Finally, we would like to humbly dedicate this work to the memory of the victims of the genocide in Gaza.
\end{acknowledgments}
